Marginally-Stable Thermal Equilibria of Rayleigh-Bénard Convection

### 2021-03-02 publication_materials/ms_rbc.tex
introduction rev 1

\begin{abstract}
A novel quasilinear form of the Boussinesq Rayleigh-B\'enard equations is derived from the classical linear stability analysis. Eigenfunction amplitudes are obtained by requiring marginal-stability. Symmetric thermally-equilibrated states are then obtained and their properties are analyzed. Such states are compared with statistically-steady 2-dimensional Rayleigh-Be\'nard convection simulations: marginally-stable thermally-equilibrated states have thinner boundary layers which scale like $\delta \sim Ra^{-1/3}$. 2-dimensional simulations are initialized with solutions to the quasilinear problem;
they do not undergo a convective transient period but turn out to be kinetically exaggerated.
\end{abstract}

@@

Rayleigh-B\'enard convection plays a foundational role in large-scale astrophysical and geophysical systems. Highly turbulent systems, such as the convective regions of stars, are not well understood. Numerical simulations provide insight into the behavior of convective systems but often times the computational expense is inhibitive. 
convection is important, particularly rbc; full rbc equations are complicated; simplified models give insight into complicated problem; chini's paper gives interesting new way to develop quasilinear approximation; jklfjakjdfskdjfajlsdjfalskdjflaskjdlfjaljsdklfjlajdlfjaklsjdklfjaljklfjakjdf skdjfajlsdjfalskdjflaskjdlfjaljsdklfjlajdlfjaklsjdklfjaljklfjakjdfskdjfajlsdjfa skdjfajlsdjfalskdjflaskjdlfjaljsdklfjlajdlfjaklsjdklfjaljklfjakjdfskdjfajlsdjfalskdjfajlsdjfalskdjflaskjdlfjaljsdklfjlajdlfjaklsjdklfjaljklfjakjdfskdjfajlsdjfal

@@

\par To derive the linearized system, we posit that each field can be represented as the sum of a background profile (denoted by $\bar{\cdot } \;$) and a perturbation function (denoted by $\cdot'$). At an arbitrary time $t_0$, suppose $\frac{\partial \bar{T}}{\partial z} (z)$ is some given function which satisfies

@@

Then the background profiles are given by

@@

\begin{figure}[h]
    \includegraphics[width=3.4in]{EV_spectrum_ol.png}
    \caption{Eigenvalue spectra for $Ra = 10^5$. The spectrum of a marginally-stable background temperature profile $\bar{T}_0$ (shown in black) has a maximum eigenvalue of 0. Given a small fixed timestep $\Delta t$, diffusion destabilizes the system, increasing its eigenvalues (shown in blue). Advection tends to stabilize the system, decreasing its eigenvalues (shown in red). It follows that there exists some proportion ($A^2$) of these two terms which yields a new marginally-stable background temperature profile $\bar{T}_1$, whose spectrum is shown in green.}
    \label{fig:iteration_spectra} 
\end{figure}

@@

\section{Properties of Thermally Equilibrated States}\label{sec:properties}
Thermally equilibrated states of this kind are solutions to a particular quasilinear form of the Rayleigh-B\'enard convection equations. For this investigation, we calculate solutions for $Ra$ in the range $10^5 - 10^9$. Solutions are symmetric about $z = 0$ by design. This, combined with the fact that we did initialize a background horizontal flow implies the absence of any nonconstant horizontal velocity perturbations.

\begin{figure}[h]
    \centering
    \includegraphics[width=3.4in]{T_profs_na.png}
    \caption{Background temperature profiles $\bar{T}$ for $Ra = 10^8$. The initial profile is given by (\ref{EQ:T0}). The equilibrated curve refers to the background temperature profile of the marginally-stable thermally-equilibrated state. The simulation profile is obtained by performing a 2-dimensional nonlinear simulation with \texttt{Dedalus} until thermal and kinetic relaxation. The simulation temperature data are horizontally- and time-averaged. The initial, equilibrated, and simulation profiles have increasingly relaxed boundary regions (respectively). The equilibrated profile exhibits prominent dips, nested alongside the boundary regions. The purpose of this feature is not well understood.}
    \label{fig:T0_profiles}
\end{figure}

\begin{figure*}[t]
    \centering
    \begin{tabular}{@{}c@{}}
        \includegraphics[width=3.4in]{flux_sup_n.png}
    \end{tabular}
    \begin{tabular}{@{}c@{}}
        \includegraphics[width=3.4in]{EV_spectra_2ra.png}
    \end{tabular}
    \caption{Heat fluxes (left) and eigenvalue spectra (right) of equilibrated states $Ra = 2 \times 10^5$ (top) and $Ra = 10^9$ (bottom). Advection profiles belong to marginally-stable modes. For low $Ra$, a single mode with $k_x = 1.5\pi$ is sufficient to oppose boundary layer diffusion and facilitate heat flux throughout the bulk of the domain. For large $Ra$, high-wavenumber modes contribute pronounced small-scale advection profiles which tightly hug the thin boundary layers. A combination of of progressively wider advection profiles is necessary to transition to the $k_x = 1.5\pi$ mode.}
    \label{fig:flux}
\end{figure*}

\begin{figure}[h]
    \centering
    \includegraphics[width=3.4in]{kx_m_ra1.png}
    \caption{Wavenumbers of marginally-stable modes of thermally equilibrated states. Markers are color-coded according to their adjacent local maxima index in the eigenvalue spectrum. For example, the spectrum corresponding to $Ra = 10^5$ has marginal wavenumbers $k_x = \pi, \, 1.5\pi$ adjacent to a presumed local maximum between these two allowed values. The $Ra = 10^9$ spectrum, shown in lower right corner of Figure \ref{fig:flux}, has three presumed local maxima, with a single marginal mode adjacent to the first maximum ($k_x = 1.5\pi$), two marginal modes adjacent to the second maximum ($k_x = 6\pi, \, 6.5\pi$), and two marginal modes adjacent to the third maximum ($k_x = 23.5\pi, \, 24\pi$). The largest wavenumbers of the green branch obey a power-law relationship with $Ra$}
    \label{fig:kx_marginals}
\end{figure}
\begin{figure}[h]
    \centering
    \includegraphics[width=3.4in]{del_ra.PNG}
    \caption{Boundary layer widths $\delta$ of marginally-stable thermal equilibria. We define the threshold of each boundary layer as the $z$-coordinate at which $\frac{\partial \bar{T}}{\partial z} = 0$, corresponding to the local extrema of the equilibrated curve in Figure \ref{fig:T0_profiles}. Plotting on a log-log scale, we find $\delta$ and $Ra$ obey a power-law relationship. We also demonstrate that $Ra^{1/3}\delta$ is approximately constant with respect to $Ra$ implying $Ra \propto  \delta^{-3}$, as predicted by \cite{Malkus}}
    \label{fig:bl_ra}
\end{figure}
\par Figure \ref{fig:T0_profiles} gives temperature profiles for $Ra = 10^8$ where the initial profile, given by (\ref{EQ:T0}), is employed at iteration 0; the equilibrated profile corresponds to the thermally relaxed state; and the simulation profile is obtained by performing a nonlinear simulation of (\ref{EQ:motion1})-(\ref{EQ:motion3}) with \texttt{Dedalus}, followed by horizontal and time averaging. The simulation curve is more diffuse than the equilibrated curve, which in turn, is more diffuse than the initial curve. Performing an eigenvalue solve by setting $\bar{T}$ equal to the simulation profile yields positive eigenvalues. Comprehensively, this suggests that thermally-equilibrated states might maximize boundary layer thickness, while subject to the marginal stability constraint.
\par The most resilient and unexpected feature of thermally equilibrated background temperature profiles are the pronounced dips adjacent to the boundary layers. These dips appear in every solution, regardless of $Ra$. Physically, they correspond to thin layers in which the temperature gradient reverses, contradicting a hypothesis of \cite{Malkus}. This counter-diffusion, which opposes overall heat transfer, is overcome by the coinciding advective flux, shown in Figure \ref{fig:flux}.

@@

\begin{figure}[h]
    \centering
    \includegraphics[width=3.4in]{del_kx_inv.png}
    \caption{For $Ra \geq 10^6$, the maximum marginally-stable wavenumber (corresponding to the green X markers in Figure \ref{fig:kx_marginals}) are inversely related to the boundary layer thicknesses $\delta$. $(\max \{ k_x \})^{-1}$ gives a minimum $x$ length scale for the perturbations, and consequently, the advection. We expect the minimum $z$ length scale to agree with the boundary layer thickness $\delta$ because otherwise the boundary layer would continue to diffuse. This is well illustrated in the lower left corner of Figure \ref{fig:flux}, where the advection profile for $k_x = 24\pi$ is tightly flanked by the surrounding boundary layers. This suggests that the minimum $x$ and $z$ length scales obey some constant ratio.}
    \label{fig:del_inv}
\end{figure}

@@

Nusselt numbers $Nu$ of various solutions are given in Figure \ref{fig:nu_vs_ra}. The marginally-stable thermally-equilibrated states and numerical simulation data exhibit power-law relationships with $Ra$. More specifically, $Ra \sim Nu^3$ which agrees with Figures \ref{fig:bl_ra}, \ref{fig:kx_marginals}, and \ref{fig:del_inv} and the scaling argument given by \cite{Malkus}. The nonlinear solutions, which maximize heat flux, are obtained \cite{Waleffe}. $Nu$ of both equilibria exceed that of simulations, which might be due to the transitional behaviors outlined by \cite{Yalniz} inhibitting heat flux.
\begin{figure*}
    \centering%
    \includegraphics[width=7.1in]{nu_ra.PNG}
    \caption{Nusselt numbers are shown for marginally-stable thermally-equilibrated states, statistically-steady direct numericals simulations performed with \texttt{Dedalus} \cite{Anders_cd}, and equilibrated nonlinear solutions, which maximize heat flux, given by \cite{Waleffe}. Both obey power-law relationships with $Ra$, with $Ra \sim Nu^3$ in agreement with Figures \ref{fig:bl_ra}, \ref{fig:kx_marginals}, and \ref{fig:del_inv} as well as \cite{Malkus}. Marginally-stable states' Nusselt numbers exceed simulation Nusselt numbers. This can be well explained by the contrasting boundary region geometries shown in Figure \ref{fig:T0_profiles}.}%
    \label{fig:nu_vs_ra}%
\end{figure*}

@@

>>>>

\begin{abstract}
A novel quasilinear form of the Boussinesq Rayleigh-B\'enard equations is derived from the classical linear stability analysis. Eigenfunction amplitudes are obtained by requiring marginal-stability. Symmetric thermally-equilibrated states are then obtained and their properties are analyzed. Such states are compared with statistically-steady 2-dimensional Rayleigh-Be\'nard convection simulations; marginally-stable thermally-equilibrated states have thinner boundary layers which scale like $\delta \sim Ra^{-1/3}$. 2-dimensional simulations are initialized with solutions to the quasilinear problem;
they do not undergo a convective transient period but turn out to be kinetically exaggerated.
\end{abstract}

@@

Rayleigh-B\'enard convection plays a foundational role in large-scale astrophysical and geophysical settings. Our understanding of highly turbulent convective systems is generally limited by the prohibitive computational costs associated with direct numerical simulations \cite{Zhu}. There exists a substantial body of work concerning the asymptotic scaling laws of such systems with no general consensous. Reduced models are particularly useful in this context because they allow us to study the underlying behavior of turbulent convection with relatively low expense. 
\par In this investigation, we derive and solve a quasilinear form of the Boussinesq Rayleigh–Bénard convection equations using a strategy similar to \cite{Chini_sw}. This is realized by solving the linearized equations and evolving a mean temperature profile according to the perturbation variables. In general, each set of perturbations has an arbitrary amplitude which we obtain by imposing marginal stability. The application of this procedure yields time-invariant solutions to the quasilinear problem with a range of expected and unexpected features. 
\par Reduced models are often valuable when supplemented by direct simulation [source?]. Nonlinear simulations perfomed with quasilinear solutions as initial conditions do not undergo a convective transient period of turbulent upwelling. Instead, the characteristic jet-like structure is somewhat present on initialization. This is practical because the convective transient period generally involves high-Reynolds number flows whose resolution is prohibitive. 

@@

\par To derive the linearized system, we posit that each field can be represented as the sum of a mean profile (denoted by $\bar{\cdot } \;$) and a perturbation function (denoted by $\cdot'$). At an arbitrary time $t_0$, suppose $\frac{\partial \bar{T}}{\partial z} (z)$ is some given function which satisfies

@@

Then the mean profiles are given by

@@

\begin{figure}[h]
    \includegraphics[width=3.4in]{EV_spectrum_ol.png}
    \caption{Eigenvalue spectra for $Ra = 10^5$. The spectrum of a marginally-stable mean temperature profile $\bar{T}_0$ (shown in black) has a maximum eigenvalue of 0. Given a small fixed timestep $\Delta t$, diffusion destabilizes the system, increasing its eigenvalues (shown in blue). Advection tends to stabilize the system, decreasing its eigenvalues (shown in red). It follows that there exists some proportion ($A^2$) of these two terms which yields a new marginally-stable mean temperature profile $\bar{T}_1$, whose spectrum is shown in green.}
    \label{fig:iteration_spectra} 
\end{figure}

@@

\section{Properties of Thermally Equilibrated States}\label{sec:properties}
Thermally equilibrated states of this kind are solutions to a particular quasilinear form of the Rayleigh-B\'enard convection equations. For this investigation, we calculate solutions for $Ra$ in the range $10^5 - 10^9$. Solutions are symmetric about $z = 0$ by design. This, combined with the fact that we did initialize a mean horizontal flow implies the absence of any nonconstant horizontal velocity perturbations. 
\par Figure \ref{fig:T0_profiles} gives temperature profiles for $Ra = 10^8$ where the initial profile, given by (\ref{EQ:T0}), is employed at iteration 0; the equilibrated profile corresponds to the thermally relaxed state; and the simulation profile is obtained by performing a nonlinear simulation of (\ref{EQ:motion1})-(\ref{EQ:motion3}) with \texttt{Dedalus}, followed by horizontal and time averaging. The simulation curve is more diffuse than the equilibrated curve, which in turn, is more diffuse than the initial curve. Performing an eigenvalue solve by setting $\bar{T}$ equal to the simulation profile yields positive eigenvalues. Comprehensively, this suggests that thermally-equilibrated states might maximize boundary layer thickness, while subject to the marginal stability constraint.
\par The most resilient and unexpected feature of thermally equilibrated mean temperature profiles are the pronounced dips adjacent to the boundary layers. These dips appear in every solution, regardless of $Ra$. Physically, they correspond to thin layers in which the temperature gradient reverses, contradicting a hypothesis of \cite{Malkus}. This counter-diffusion, which opposes overall heat transfer, is overcome by the coinciding advective flux, shown in Figure \ref{fig:flux}. 
\begin{figure}[h]
    \centering
    \includegraphics[width=3.4in]{T_profs_na.png}
    \caption{mean temperature profiles $\bar{T}$ for $Ra = 10^8$. The initial profile is given by (\ref{EQ:T0}). The equilibrated curve refers to the mean temperature profile of the marginally-stable thermally-equilibrated state. The simulation profile is obtained by performing a 2-dimensional nonlinear simulation with \texttt{Dedalus} until thermal and kinetic relaxation. The simulation temperature data are horizontally- and time-averaged. The initial, equilibrated, and simulation profiles have increasingly relaxed boundary regions (respectively). The equilibrated profile exhibits prominent dips, nested alongside the boundary regions. The purpose of this feature is not well understood.}
    \label{fig:T0_profiles}
\end{figure}

\begin{figure}[h]
    \centering
    \includegraphics[width=3.4in]{kx_m_ra1.png}
    \caption{Wavenumbers of marginally-stable modes of thermally equilibrated states. Markers are color-coded according to their adjacent local maxima index in the eigenvalue spectrum. For example, the spectrum corresponding to $Ra = 10^5$ has marginal wavenumbers $k_x = \pi, \, 1.5\pi$ adjacent to a presumed local maximum between these two allowed values. The $Ra = 10^9$ spectrum, shown in lower right corner of Figure \ref{fig:flux}, has three presumed local maxima, with a single marginal mode adjacent to the first maximum ($k_x = 1.5\pi$), two marginal modes adjacent to the second maximum ($k_x = 6\pi, \, 6.5\pi$), and two marginal modes adjacent to the third maximum ($k_x = 23.5\pi, \, 24\pi$). The largest wavenumbers of the green branch obey a power-law relationship with $Ra$}
    \label{fig:kx_marginals}
\end{figure}
\begin{figure}[h]
    \centering
    \includegraphics[width=3.4in]{del_kx_inv.png}
    \caption{For $Ra \geq 10^6$, the maximum marginally-stable wavenumber (corresponding to the green X markers in Figure \ref{fig:kx_marginals}) are inversely related to the boundary layer thicknesses $\delta$. $(\max \{ k_x \})^{-1}$ gives a minimum $x$ length scale for the perturbations, and consequently, the advection. We expect the minimum $z$ length scale to agree with the boundary layer thickness $\delta$ because otherwise the boundary layer would continue to diffuse. This is well illustrated in the lower left corner of Figure \ref{fig:flux}, where the advection profile for $k_x = 24\pi$ is tightly flanked by the surrounding boundary layers. This suggests that the minimum $x$ and $z$ length scales obey some constant ratio.}
    \label{fig:del_inv}
\end{figure}

\begin{figure}[h]
    \centering
    \includegraphics[width=3.4in]{del_ra.PNG}
    \caption{Boundary layer widths $\delta$ of marginally-stable thermal equilibria. We define the threshold of each boundary layer as the $z$-coordinate at which $\frac{\partial \bar{T}}{\partial z} = 0$, corresponding to the local extrema of the equilibrated curve in Figure \ref{fig:T0_profiles}. Plotting on a log-log scale, we find $\delta$ and $Ra$ obey a power-law relationship. We also demonstrate that $Ra^{1/3}\delta$ is approximately constant with respect to $Ra$ implying $Ra \propto  \delta^{-3}$, as predicted by \cite{Malkus}}
    \label{fig:bl_ra}
\end{figure}

\begin{figure*}
    \centering
    \begin{tabular}{@{}c@{}}
        \includegraphics[width=3.4in]{flux_sup_n.png}
    \end{tabular}
    \begin{tabular}{@{}c@{}}
        \includegraphics[width=3.4in]{EV_spectra_2ra.png}
    \end{tabular}
    \caption{Heat fluxes (left) and eigenvalue spectra (right) of equilibrated states $Ra = 2 \times 10^5$ (top) and $Ra = 10^9$ (bottom). Advection profiles belong to marginally-stable modes. For low $Ra$, a single mode with $k_x = 1.5\pi$ is sufficient to oppose boundary layer diffusion and facilitate heat flux throughout the bulk of the domain. For large $Ra$, high-wavenumber modes contribute pronounced small-scale advection profiles which tightly hug the thin boundary layers. A combination of of progressively wider advection profiles is necessary to transition to the $k_x = 1.5\pi$ mode.}
    \label{fig:flux}
    \vspace{0.4in}
    \includegraphics[width=7.1in]{nu_ra.PNG}
    \caption{Nusselt numbers are shown for marginally-stable thermally-equilibrated states, statistically-steady direct numericals simulations performed with \texttt{Dedalus} \cite{Anders_cd}, and equilibrated nonlinear solutions, which maximize heat flux, given by \cite{Waleffe}. Both obey power-law relationships with $Ra$, with $Ra \sim Nu^3$ in agreement with Figures \ref{fig:bl_ra}, \ref{fig:kx_marginals}, and \ref{fig:del_inv} as well as \cite{Malkus}. Marginally-stable states' Nusselt numbers exceed simulation Nusselt numbers. This can be well explained by the contrasting boundary region geometries shown in Figure \ref{fig:T0_profiles}.}%
    \label{fig:nu_vs_ra}%
\end{figure*}

@@

@@

Nusselt numbers $Nu$ of various solutions are given in Figure \ref{fig:nu_vs_ra}. The marginally-stable thermally-equilibrated states and numerical simulation data exhibit power-law relationships with $Ra$. More specifically, $Ra \sim Nu^3$ which agrees with Figures \ref{fig:bl_ra}, \ref{fig:kx_marginals}, and \ref{fig:del_inv} and the scaling argument given by \cite{Malkus}. The nonlinear solutions, which maximize heat flux, are obtained \cite{Waleffe}. $Nu$ of both equilibria exceed that of simulations, which might be due to the transitional behaviors outlined by \cite{Yalniz} inhibitting heat flux.

@@

\bibitem{Zhu} Xiaojue Zhu, Varghese Mathai, Richard J. A. M. Stevens, Roberto Verzicco, and Detlef Lohse (2018). Transition to the Ultimate Regime in Two-Dimensional Rayleigh-Bénard Convection. Phys. Rev. Lett. 120, 144502
